Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-22 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
updated theory discussion, in part. dcs

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General:
The detection of gravitational waves produced by coalescing compact binaries
allows us to explore the dynamical, nonlinear regime of general relativity
and to constrain modifications to it.
%
Some of these gravitational-wave events have sufficiently high
signal-to-noise ratio in the merger, allowing us to probe the relaxation of the remnant black
hole to its final, stationary state --- the so-called black-hole ringdown, which is
characterized by a set of quasinormal modes.
% Main question:
Can we use the ringdown to constrain deviations to general relativity,
as predicted by several of its contenders?
% What we do:
Here, we  address this question by using an inspiral-merger-ringdown waveform
model in the effective-one-body formalism, augmented with a parametrization of
the ringdown based on an expansion in the final black hole's spin.
%
We give a prescription on how to include in this waveform model the quasinormal mode frequencies
calculated on a theory-by-theory basis. In particular, we focus on gravity theories that deviate
from general relativity through higher-order curvature corrections, namely,
Einstein-dilaton-Gauss-Bonnet, dynamical Chern-Simons, cubic- and quartic-order
effective-field-theories of general relativity.
%
% Main conclusion:
We confront the resultant waveform models against the two loudest
gravitational-wave ringdown signals, so far, notably, GW150914 and GW200129.
%
We find that while the Einstein-dilaton-Gauss-Bonnet theory cannot be
constrained with these events, we can place upper bounds on the coupling
constants of cubic- ($\ell_{\rm cEFT} \leqslant 38.2$~km) and quartic-order
($\ell_{\rm qEFT} \leqslant 51.3$~km) effective-field-theories of general
relativity, and of dynamical Chern-Simons gravity ($\ell_{\rm dCS} \leqslant 38.7$~km).
%
The latter result is a concrete example of a theory presently unconstrained by
inspiral-only analyses which, however, can be constrained by merger-ringdown
studies with current gravitational-wave data.
%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

@@

Since the first detection of gravitational waves (GWs) from a binary black-hole (BBH) merger in 2015~\cite{LIGOScientific:2016aoc},
the LIGO~\cite{LIGOScientific:2014pky} and Virgo~\cite{VIRGO:2014yos} detectors have observed about 90 GWs signals \cite{LIGOScientific:2021djp} from mergers
of BHs, neutron stars (NSs) \cite{TheLIGOScientific:2017qsa,LIGOScientific:2020aai} and their mixture \cite{LIGOScientific:2021qlt}. These results have been confirmed
by independent analyses, which have also identified a few new GW signals~\cite{Nitz:2018imz,Nitz:2019hdf,Venumadhav:2019lyq,Zackay:2019btq,Nitz:2021zwj,Olsen:2022pin}.

@@

This theory belongs to the class of scalar-Gauss-Bonnet theories, which are
described by the action
%
\begin{align} \label{eq:action_sgb}
    S_{\rm EdGB} = \frac{1}{16 \pi}
    \int \dV
    \left[
    R - \tfrac{1}{2}(\nabla \varphi)^2
    + \tfrac{1}{4} \ell^{2}_{\rm EdGB} f(\varphi) \, \mathcal{G}
    \right]\,,
    \nonumber \\
\end{align}
%
where $g \equiv \textrm{det}(g_{\mu \nu})$ is the metric determinant, $R$ is the Ricci
scalar, $\varphi$ is a dynamical scalar field that couples to the
Gauss-Bonnet invariant
%
$\mathcal{G} =
R^{\mu\nu\rho\sigma}R_{\mu\nu\rho\sigma}
- 4 R^{\mu\nu}R_{\mu\nu}
+ R^2$,
%
and $R_{\mu\nu\rho\sigma}$ and $R_{\mu\nu}$ are the Riemann and Ricci tensors respectively.
%
By itself, $\int \dV \mathcal{G}$ is a boundary term in four dimensions
and hence does not contribute to the field equations~\cite{Myers:1987yn}.
%
However,  when coupled to $\varphi$, it can contribute to the field equation
through the coupling function $f(\varphi)$. The strength of the coupling is
set by $\ell_{\rm GB}$, with dimensions of length.

Different subclasses of this theory are determined by the function $f(\varphi)$
and can be divided into two classes based on the properties of their BH
solutions.
%
In the first class, the first derivative of the coupling function $f'(\varphi)
\equiv \dd f  / \dd \varphi$ is always nonzero and BHs are known to always
support scalar hair.
%
This is the case of Einstein-dilaton-Gauss-Bonnet (EdGB) gravity, for which
$f(\varphi) = \exp(\varphi)$~\cite{Kanti:1995vq}.
%
In the second class, $f'(\varphi) = 0$ can vanish for some constant $\varphi_0$.
%
In this case, the theory admits the same stationary, asymptotically flat BH
solutions as GR and those of scalarized
BHs~\cite{Doneva:2017bvd,Silva:2017uqg,Macedo:2019sem,Dima:2020yac,Herdeiro:2020wei,Berti:2020kgk}.
%
Examples include the Gaussian $f(\varphi) \propto \exp(-\varphi^2)$~\cite{Doneva:2017bvd} and
the quadratic $f(\varphi) \propto \varphi^2$~\cite{Silva:2017uqg} coupling functions.
BHs in EdGB gravity have scalar hair, to which we can associate a (monopole) scalar charge $Q$,
related to the asymptotic $r^{-1}$ fall-off of the scalar field, where $r$ is the distance from the BH.
%
This charge is not an independent parameter, and depends on the BH's mass and spin,
thus being a ``secondary hair''~\cite{Coleman:1991ku,Herdeiro:2015waa}.
%
Since the scalar field is sourced by a curvature scalar, $Q$ is larger (smaller),
the smaller (larger) the BH mass is.

These properties are not mere theoretical curiosities; they have important observational consequences.
%
First, the presence of the scalar charge implies that when in binaries, BHs can source
scalar-dipole radiation (see, e.g., Refs.~\cite{Yagi:2011xp,Julie:2019sab,Shiralilou:2020gah,Shiralilou:2021mfl,Julie:2022huo})
which affects the GW phase at $-1$PN order (relative to the dominant quadrupolar GR
contribution), with magnitude proportional to the difference between the
charges of binary components. This makes EdGB gravity testable with present GW
observations of compact binaries where at least if one component is a BH.
%
More specifically, Ref.~\cite{Lyu:2022gdr} placed the bound
%
$\ell_{\rm EdGB} \leqslant 1.33$~km
%
using the NSBH binaries GW200105 and
GW200115~\cite{LIGOScientific:2021qlt} while Refs.~\cite{Nair:2019iur,Perkins:2021mhb}, obtained
%
$\ell_{\rm EdGB} \leqslant 1.7$~km,
%
by stacking the posteriors of $\ell_{\rm EdGB}$ from a selection of BBHs
from the GWTC-1 and GWTC-2 catalogs~\cite{LIGOScientific:2018mvr,LIGOScientific:2020ibl}~\footnote{These bounds
are, strictly speaking, valid only when the scalar field $\varphi$ is small (i.e., $\varphi \ll 1$), and
take into account only the leading-order scalar field contribution arising from the dilatonic coupling.}.
Secondly, the scalar field does influence the response of BHs to linearized perturbations and thus
affects the BH QNM spectra.
%
The coupling between scalar field and the Gauss-Bonnet invariant, which
result in a coupling between scalar perturbations and gravitational
perturbations of polar parity can, for example break the equivalence
between the QNM spectra of polar and axial gravitational perturbations~\cite{Glampedakis:2017rar,Lenzi:2021njy}
(i.e., the isospectrality~\cite{Chandrasekhar:1985kt}) for Schwarzschild BHs in GR.
%

The QNMs for nonrotating BHs in EdGB gravity were first computed in
Refs.~\cite{Pani:2009wy,Blazquez-Salcedo:2016enn} and were
extended to leading-order in BH spin in Ref.~\cite{Pierini:2021jxd}.

@@

This theory is described by the action~\cite{Jackiw:2003pm,Alexander:2009tp},
%
\begin{align} \label{eq:action_dcs}
    S_{\rm dCS} = \frac{1}{16 \pi}
    \int \dV
    \left[
    R - \tfrac{1}{2}(\nabla \vartheta)^2
    + \tfrac{1}{4} \ell^{2}_{\rm dCS} \, \vartheta \, {}^{*}RR
    \right],
    \nonumber \\
\end{align}
%
where $\vartheta$ is a pseudoscalar field which couples to the Pontryagin
density
${}^{*}RR = {}^{*}R_{\mu\nu}{}^{\rho\sigma} R^{\nu\mu}{}_{\rho\sigma}$,
%
where ${}^{*}R^{\mu\nu\rho\sigma}$ is the dual of the Riemann tensor
defined as
%
${}^{*}R^{\mu\nu\rho\sigma} =
\epsilon^{\mu\nu}{}_{\gamma\delta}
R^{\gamma\delta\rho\sigma} / 2$,
%
and $\epsilon_{\mu\nu\gamma\delta}$ is the Levi-Civita tensor.
%
As in scalar-Gauss-Bonnet theories, $\int \dV {}^{*}RR$ by itself is a boundary
term~\cite{Jackiw:2003pm} that does not contribute to the field equations in
four-dimensions.
%
However, a coupling with $\vartheta$ can modify the field equations.
%
The coupling strength is set by the coupling constant $\ell_{\rm dCS}$, with
dimensions of length.

The theory admits as a solution, the garden-variety Schwarzschild BH of GR. This
is not the case when rotation is included and the Kerr metric is not a solution of
the theory~\cite{Jackiw:2003pm}.
%
These rotating BHs support a scalar field which fall off as
$r^{-2}$ asymptotically, to which we can associate a scalar dipole
charge~\cite{Yunes:2009hc,Konno:2009kg}.
%
As a consequence, the leading-scalar radiative channel of BBH to lose energy is quadrupolar,
which affects the GW phase at 2PN order~\cite{Yagi:2011xp,Alexander:2017jmt}.
%
Deviations from GR at this PN order are constrained with present GW observations~\cite{LIGOScientific:2021sio}
\ab{only at the level of $\sim 50\%$ (see the constraint on the 2PN parameter $\varphi_4$ in Fig.XXX of
Ref.~\cite{LIGOScientific:2021sio}). Thus, dCS gravity is poorly constrained by
analyses that rely on the inspiral portion of GWs when using current BBH observations are used~\cite{Nair:2019iur,Perkins:2021mhb}.}
%
Nonetheless, the theory has been constrained in Ref.~\cite{Silva:2020acr}, which found
%
$\ell_{\rm dCS} \leqslant 8.5$~km,
%
by folding data from the X-ray observations of the isolated NS
PSR~J0030+0451~\cite{Lommen:2000yt,NANOGrav:2017wvv} by
NICER~\cite{Riley:2019yda,Miller:2019cac} and from the GW observation of the binary
NS GW170817~\cite{TheLIGOScientific:2017qsa,LIGOScientific:2018cki}.

The QNMs of the Schwarzschild BH in dCS gravity were studied in
Refs.~\cite{Yunes:2007ss,Cardoso:2009pk,Molina:2010fb}, which found that scalar
perturbations couple to gravitational perturbations of axial parity, in contrast
with EdGB gravity.
%
This leads to a breakdown of isospectrality. The QNM spectra of slowly-rotating BHs in dCS gravity was
studied in Ref.~\cite{Srivastava:2021imr,Wagle:2021tam}.
%
They were also extracted from numerical-relativity simulations of head-on BH collisions
in Ref.~\cite{Okounkova:2019dfo}.

@@

More specifically, we follow the notation of
Refs.~\cite{Cano:2020cao,Cano:2021myl} and consider up to dimension-eight
operators ($n=4$),
%
\begin{subequations}
\label{eq:action_mods}
\begin{align}
    L^{(6)} &= \lame R_{\mu\nu}{}^{\rho\sigma} R_{\rho\sigma}{}^{\gamma\delta} R_{\gamma \delta}{}^{\mu\nu}
    + \lamo R_{\mu\nu}{}^{\rho\sigma} R_{\rho\sigma}{}^{\gamma\delta} \tilde{R}_{\gamma \delta}{}^{\mu\nu}\,,
    \label{eq:s_6}
    \\
    L^{(8)} &= \varepsilon_{1} \mathcal{C}^2
    + \varepsilon_{2} \tilde{\mathcal{C}}^{2}
    + \varepsilon_{3} \mathcal{C} \tilde{\mathcal{C}}\,,
\label{eq:s_8}
\end{align}
\end{subequations}
%
where ${\cal C} = R_{\mu\nu\rho\sigma} R^{\mu\nu\rho\sigma}$,
$\tilde{\cal C} = R_{\mu\nu\gamma\delta} \epsilon^{\mu\nu}{}_{\rho\sigma} R^{\rho\sigma\gamma\delta}$,
and $(\lambda_{\rm o, e},\varepsilon_{i})$, $i=$~1, 2, 3, are dimensionless parameters.
%
Due to the large number of free parameters in this theory, we focus on a subset of the parameter space.
%
In particular, we consider dimension-six and dimension-eight operators separately.
%
In addition, in the dimension-six case we further assume that $\lame = \lamo = 1$,
leaving us with \redsout{$\ell =$} $\ell_{\rm cEFT}$ as our single free parameter.
%
Similarly, in the dimension-eight case, we set $\varepsilon_{1} = 1$ and $\varepsilon_{2} = \varepsilon_{3} = 0$,
as done in Ref.~\cite{Sennett:2019bpc}.
%
This leaves us with \redsout{$\ell =$} $\ell_{\rm qEFT}$ as our single free parameter.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General:
The detection of gravitational waves produced by coalescing compact binaries
allows us to explore the dynamical, nonlinear regime of general relativity
and to constrain modifications to it.
%
Some of these gravitational-wave events have sufficiently high
signal-to-noise ratio in the merger, allowing us to probe the relaxation of the remnant black
hole to its final, stationary state --- the so-called black-hole ringdown, which is
characterized by a set of quasinormal modes.
% Main question:
Can we use the ringdown to constrain deviations to general relativity,
as predicted by several of its contenders?
% What we do:
Here, we  address this question by using an inspiral-merger-ringdown waveform
model in the effective-one-body formalism, augmented with a parametrization of
the ringdown based on an expansion in the final black hole's spin.
%
We give a prescription on how to include in this waveform model the quasinormal mode frequencies
calculated on a theory-by-theory basis. In particular, we focus on gravity theories that deviate
from general relativity through higher-order curvature corrections, namely,
Einstein-dilaton-Gauss-Bonnet, dynamical Chern-Simons, cubic- and quartic-order
effective-field-theories of general relativity.
%
% Main conclusion:
We confront the resultant waveform models against the two loudest
gravitational-wave ringdown signals observed so far, notably, GW150914 and GW200129.
%
We find that while the Einstein-dilaton-Gauss-Bonnet theory cannot be
constrained with these events, we can place upper bounds on the coupling
constants of cubic- ($\ell_{\rm cEFT} \leqslant 38.2$~km) and quartic-order
($\ell_{\rm qEFT} \leqslant 51.3$~km) effective-field-theories of general
relativity, and of dynamical Chern-Simons gravity ($\ell_{\rm dCS} \leqslant 38.7$~km).
%
The latter result is a concrete example of a theory presently unconstrained by
inspiral-only analyses which, however, can be constrained by merger-ringdown
studies with current gravitational-wave data.
%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

@@

Since the first detection of gravitational waves (GWs) from a binary black-hole (BBH) merger in 2015~\cite{LIGOScientific:2016aoc},
the LIGO~\cite{LIGOScientific:2014pky} and Virgo~\cite{VIRGO:2014yos} detectors have observed about 90 GWs signals \cite{LIGOScientific:2021djp} from mergers
of BHs, neutron stars (NSs) \cite{TheLIGOScientific:2017qsa,LIGOScientific:2018cki,LIGOScientific:2020aai} and their mixture \cite{LIGOScientific:2021qlt}. These results have been confirmed
by independent analyses, which have also identified a few new GW signals~\cite{Nitz:2018imz,Nitz:2019hdf,Venumadhav:2019lyq,Zackay:2019btq,Nitz:2021zwj,Olsen:2022pin}.

@@

This theory belongs to the class of scalar-Gauss-Bonnet theories, which are
described by the action
%
\begin{align} \label{eq:action_sgb}
    S_{\rm EdGB} = \frac{1}{16 \pi}
    \int \dV
    \left[
    R - \tfrac{1}{2}(\partial \varphi)^2
    + \tfrac{1}{4} \ell^{2}_{\rm EdGB} f(\varphi) \, \mathcal{G}
    \right],
    \nonumber \\
\end{align}
%
where $g \equiv \textrm{det}(g_{\mu \nu})$ is the metric determinant, $R$ is the Ricci
scalar, $\varphi$ is a dynamical scalar field, with kinetic term $(\partial \varphi)^2 = g^{\mu\nu} \partial_{\mu} \varphi \partial_{\nu} \varphi$,
which couples to the Gauss-Bonnet invariant
%
$\mathcal{G} =
R^{\mu\nu\rho\sigma}R_{\mu\nu\rho\sigma}
- 4 R^{\mu\nu}R_{\mu\nu}
+ R^2$,
%
and $R_{\mu\nu\rho\sigma}$ and $R_{\mu\nu}$ are the Riemann and Ricci tensors respectively.
%
By itself, $\int \dV \mathcal{G}$ is a boundary term in four dimensions
and hence does not contribute to the field equations~\cite{Myers:1987yn}.
%
However,  when coupled to $\varphi$, it can contribute to the field equation
through the coupling function $f(\varphi)$. The strength of the coupling is
set by $\ell_{\rm GB}$, with dimensions of length.

Different subclasses of this theory are determined by the function $f(\varphi)$
and can be divided into two classes based on the properties of their BH
solutions.
%
In the first class, the first derivative of the coupling function $f'(\varphi)
\equiv \dd f  / \dd \varphi$ is always nonzero and BHs are known to always
support scalar hair.
%
This is the case of Einstein-dilaton-Gauss-Bonnet (EdGB) gravity, for which
$f(\varphi) = \exp(\varphi)$~\cite{Kanti:1995vq}.
%
In the second class, $f'(\varphi) = 0$ can vanish for some constant $\varphi_0$.
%
In this case, the theory admits the same stationary, asymptotically flat BH
solutions as GR~\cite{Silva:2017uqg} and those of scalarized
BHs~\cite{Doneva:2017bvd,Silva:2017uqg,Macedo:2019sem,Dima:2020yac,Herdeiro:2020wei,Berti:2020kgk}.
%
Examples include Gaussian $f(\varphi) \propto \exp(-\varphi^2)$~\cite{Doneva:2017bvd} and
the quadratic $f(\varphi) \propto \varphi^2$~\cite{Silva:2017uqg} coupling functions.

BHs in EdGB gravity have scalar hair, to which we can associate a (monopole) scalar charge $Q$,
related to the asymptotic $r^{-1}$ fall-off of the scalar field, where $r$ is the distance from the BH.
%
This charge is not an independent parameter, and depends on the BH's mass and spin,
thus being a ``secondary hair''~\cite{Coleman:1991ku,Kanti:1995vq,Herdeiro:2015waa}.
%
Since the scalar field is sourced by a curvature scalar, $Q$ is larger (smaller),
the smaller (larger) the BH mass is.

These properties are not mere theoretical curiosities; they have important observational consequences.
%
First, the presence of the scalar charge implies that when in binaries, BHs can source
scalar-dipole radiation (see, e.g., Refs.~\cite{Yagi:2011xp,Julie:2019sab,Shiralilou:2020gah,Shiralilou:2021mfl,Julie:2022huo})
which affects the GW phase at $-1$PN order (relative to the dominant quadrupolar GR
contribution), with magnitude proportional to the difference between the
charges of binary components. This makes EdGB gravity testable with present GW
observations of compact binaries where at least one component is a BH.
%
More specifically, Ref.~\cite{Lyu:2022gdr} placed the bound
%
$\ell_{\rm EdGB} \leqslant 1.33$~km
%
using the NSBH binaries GW200105 and
GW200115~\cite{LIGOScientific:2021qlt} while Refs.~\cite{Nair:2019iur,Perkins:2021mhb}, obtained
%
$\ell_{\rm EdGB} \leqslant 1.7$~km,
%
by stacking the posteriors of $\ell_{\rm EdGB}$ from a selection of BBHs
from the GWTC-1 and GWTC-2 catalogs~\cite{LIGOScientific:2018mvr,LIGOScientific:2020ibl}\footnote{These bounds
are, strictly speaking, valid only when the scalar field $\varphi$ is small (i.e., $\varphi \ll 1$), and
take into account only the leading-order scalar field contribution arising from the dilatonic coupling.}.
Secondly, the scalar field influences the response of BHs to linearized perturbations and thus
affects the BH's QNM spectra.
%
The coupling between scalar field and the Gauss-Bonnet invariant, which
result in a coupling between scalar perturbations and gravitational
perturbations of polar parity can, for example break the equivalence
between the QNM spectra of polar and axial gravitational perturbations~\cite{Glampedakis:2017rar,Lenzi:2021njy}
(i.e., the isospectrality~\cite{Chandrasekhar:1985kt}) for Schwarzschild BHs in GR.

Using BH perturbation theory, the QNMs for nonrotating BHs in EdGB gravity
were first computed in Refs.~\cite{Pani:2009wy,Blazquez-Salcedo:2016enn}
and were extended to leading-order in BH spin in Ref.~\cite{Pierini:2021jxd}.
%
See also Refs.~\cite{Witek:2018dmd,Okounkova:2019zep} for NR studies.

@@

This theory is described by the action~\cite{Jackiw:2003pm,Alexander:2009tp},
%
\begin{align} \label{eq:action_dcs}
    S_{\rm dCS} = \frac{1}{16 \pi}
    \int \dV
    \left[
    R - \tfrac{1}{2}(\partial \vartheta)^2
    + \tfrac{1}{4} \ell^{2}_{\rm dCS} \, \vartheta \, {}^{*}RR
    \right],
    \nonumber \\
\end{align}
%
where $\vartheta$ is a pseudoscalar field which couples to the Pontryagin
density
${}^{*}RR = {}^{*}R_{\mu\nu}{}^{\rho\sigma} R^{\nu\mu}{}_{\rho\sigma}$,
%
where ${}^{*}R^{\mu\nu\rho\sigma}$ is the dual of the Riemann tensor
defined as
%
${}^{*}R^{\mu\nu\rho\sigma} =
\epsilon^{\mu\nu}{}_{\gamma\delta}
R^{\gamma\delta\rho\sigma} / 2$,
%
and $\epsilon_{\mu\nu\gamma\delta}$ is the Levi-Civita tensor.
%
As in scalar-Gauss-Bonnet theories, $\int \dV {}^{*}RR$ by itself is a boundary
term~\cite{Jackiw:2003pm} that does not contribute to the field equations in
four-dimensions.
%
The Pontryagin density can, however, modify the field equation when coupled to $\vartheta$;
the strength of the coupling set by $\ell_{\rm dCS}$, with dimensions of length.
%
% However, a coupling with $\vartheta$ can modify the field equations.
% %
% The coupling strength is set by the coupling constant $\ell_{\rm dCS}$, with
% dimensions of length.

The theory admits as a solution, the garden-variety Schwarzschild BH of GR.
This is not the case when rotation is included and the Kerr metric is not a
solution of the theory~\cite{Jackiw:2003pm}.
%
These rotating BHs support a scalar field which fall off as
$r^{-2}$ asymptotically, to which we can associate a scalar dipole
charge~\cite{Yunes:2009hc,Konno:2009kg} and
%
the leading-order modification to the GW phase enters at 2PN~\cite{Yagi:2011xp}.
%
Deviations from GR at this PN order are constrained with present GW observations~\cite{LIGOScientific:2021sio}
only at the level of $\sim 50\%$ (see the constraint on the 2PN parameter $\varphi_4$ in Fig.~6 of
Ref.~\cite{LIGOScientific:2021sio}). Thus, so far, dCS gravity has remained
unconstrained by works which use only on the inspiral portion of BBH GW signals~\cite{Nair:2019iur,Perkins:2021mhb}.
%
Nonetheless, the theory has been constrained in Ref.~\cite{Silva:2020acr}, which found
%
$\ell_{\rm dCS} \leqslant 8.5$~km,
%
by folding data from the X-ray observations of the isolated NS
PSR~J0030+0451~\cite{Lommen:2000yt,NANOGrav:2017wvv} by
NICER~\cite{Riley:2019yda,Miller:2019cac} and from the GW observation of the binary
NS GW170817~\cite{TheLIGOScientific:2017qsa,LIGOScientific:2018cki}, using
equation-of-state independent relations between NS moment of inertia and tidal deformability~\cite{Yagi:2013bca,Yagi:2013awa,Gupta:2017vsl}.

The QNMs of the Schwarzschild BH in dCS gravity were studied in
Refs.~\cite{Yunes:2007ss,Cardoso:2009pk,Molina:2010fb}, which found that scalar
perturbations couple to gravitational perturbations of axial parity, in contrast
with EdGB gravity.
%
This leads to a breakdown of isospectrality. The QNM spectra of slowly-rotating BHs in dCS gravity was
studied in Ref.~\cite{Srivastava:2021imr,Wagle:2021tam}.
%
They were also extracted from NR simulations of BH head-on collisions in Ref.~\cite{Okounkova:2019dfo}.

@@

More specifically, we follow the notation of
Refs.~\cite{Cano:2020cao,Cano:2021myl} and consider up to dimension-eight
operators ($n=4$),
%
\begin{subequations}
\label{eq:action_mods}
\begin{align}
    L^{(6)} &= \lame R_{\mu\nu}{}^{\rho\sigma} R_{\rho\sigma}{}^{\gamma\delta} R_{\gamma \delta}{}^{\mu\nu}
    + \lamo R_{\mu\nu}{}^{\rho\sigma} R_{\rho\sigma}{}^{\gamma\delta} \tilde{R}_{\gamma \delta}{}^{\mu\nu}\,,
    \label{eq:s_6}
    \\
    L^{(8)} &= \varepsilon_{1} \mathcal{C}^2
    + \varepsilon_{2} \tilde{\mathcal{C}}^{2}
    + \varepsilon_{3} \mathcal{C} \tilde{\mathcal{C}}\,,
\label{eq:s_8}
\end{align}
\end{subequations}
%
where ${\cal C} = R_{\mu\nu\rho\sigma} R^{\mu\nu\rho\sigma}$,
$\tilde{\cal C} = R_{\mu\nu\gamma\delta} \epsilon^{\mu\nu}{}_{\rho\sigma} R^{\rho\sigma\gamma\delta}$,
and $(\lambda_{\rm o, e},\varepsilon_{i})$, $i=$~1, 2, 3, are dimensionless parameters.
%
Due to the large number of free parameters in this theory, we focus on a subset of the parameter space.
%
In particular, we consider dimension-six and dimension-eight operators separately.
%
In addition, in the dimension-six case we further assume that $\lame = \lamo = 1$,
leaving us with $\ell_{\rm cEFT}$ as our single free parameter.
%
Similarly, in the dimension-eight case, we set $\varepsilon_{1} = 1$ and $\varepsilon_{2} = \varepsilon_{3} = 0$,
as done in Ref.~\cite{Sennett:2019bpc}.
%
This leaves us with $\ell_{\rm qEFT}$ as our single free parameter.
